Appendix 3.B: Preparing OECD industry data

### 2023-02-10 PhD_pages/Paper_1_within_country_ineq/data_appendix/1_B_prepare_WID.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

@@

df_wid = pd.read_csv("data/original/wid.csv")

@@

The dataframe looks like this:
```{python}
df_wid.head(50)
```
:::

@@

```{python}
# Gini as index
df_wid["Gini"] = df_wid['Gini.coefficient'] * 100

@@

df_wid = df_wid.rename(columns={
    'P90.P100...share.of.the.top.10.':'Top_10_share',
    'P0.P50...share.of.the.bottom.50.':'Bottom_50_share',
    'P90...income.threshold.of.the.top.10.':'P90',
    'P10...income.threshold':'P10'
  })

keep_cols = [
  "Year",
  "Entity",
  "Gini",
  "Top_10_share",
  "Bottom_50_share"
]

@@

>>>>

The data was prepared in three separate files according to three income concepts: post_tax disposable income, post_tax national income and pretax national income.

These we append into a single file.

@@

inc_concepts = [
  'posttax_dis',
  'posttax_nat',
  'pretax'
]

# filepaths = [f'data/original/wid_{inc_concept}.csv' for inc_concept in inc_concepts]

filepaths = [f'https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_{inc_concept}.csv' for inc_concept in inc_concepts]

wid_dfs = [pd.read_csv(filepath) for filepath in filepaths]

df_wid = pd.concat(wid_dfs, keys=inc_concepts)

# Tidy up indexes
df_wid = df_wid.reset_index()

df_wid = df_wid.drop('level_1', axis=1)

df_wid = df_wid\
  .rename(columns={"level_0": "inc_concept"})

@@

The resulting dataframe looks like this:
```{python}
df_wid.head(50)
```
:::

@@

```{python}
# Gini as index
df_wid["Gini"] = df_wid['Gini coefficient'] * 100

@@

df_wid = df_wid.rename(columns={
    'P90-P100 - share of the top 10%':'Top_10_share',
    'P0-P50 - share of the bottom 50%':'Bottom_50_share',
    'P99-P100 - share of the top 1%':'Top_1_share'
  })

keep_cols = [
  "Year",
  "Entity",
  "Gini",
  "Top_10_share",
  "Bottom_50_share",
  "inc_concept"
]

@@

::: {.callout-note collapse="true" appearance="minimal"}
### Pivot the data wider by income concept

df_wid = df_wid.set_index(['Year','Entity','inc_concept']).unstack()

df_wid.columns = df_wid.columns.map(lambda x: '{}_{}'.format(x[0], x[1]))

df_wid = df_wid.reset_index()

df_wid = df_wid.sort_values(by=['Entity', 'Year'])

### 2023-02-10 PhD_pages/Paper_1_within_country_ineq/paper1.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

@@

    2) Hampered by the available data
        - **Coverage** – for instance no surveys.
        - **Comparability of data** – Some secondary databases with global coverage mixes very heterogenous datapoints (WIID) (Find Atkinson and Brandolini ref from Handbook...). Those with a greater degree of harmonization tend to lack global coverage (LIS, EU-SILC, SEDLAC, until recently PIP)
        - **Some trade off between the two is inevitable** – There is no global survey of incomes. Countries run surveys based on their own needs and capacities. As such they differ. Saying something global means accepting a degree of heterogenity in the data. Similarly, definitions of income for tax purposes vary, impacting the comparability of tax data.

    3) Contradictory global views are presented by different global databases.
        - The data situation has improved drastically: thanks to harmonized databases with global coverage. In particular:
            - PIP – recently expanded to include high income countries
            - WID – Transition from WTID, and the DINA methodology, global coverage.
            - LIS: some data from middle and low income countries.
        - But the 'take home' messages from the institutions producing these databases offer very contradictory views.
            - "In recent decades, income inequality 
has increased in nearly all countries, 
but at different speeds, suggesting 
that institutions and policies matter in shaping inequality." (WIR 2018)
            - "[Across a range of approaches] aver-
age  inequality  within  countries  increased   in  the  1990s,  but  has  been  falling  in  recent years according to most specifications." (World Bank, 2016)

>>>>

# Outline

- We need a global view on inequality

- Though there is much data, we don't have a clear global view.    
  [UN summary of inequality trends](https://www.un.org/en/un75/inequality-bridging-divide)

- A global view is hampered by many factors:
    - In research and commentary, there is a disportionate focus high-income countries, in particular the US

    - The extent and quality of available data
        - Coverage
        - Quality and comarabilty of data
        - Some trade off between the two is inevitable

    - **The major global inequality databases occupy different 'worlds' of measurement**

    It is the latter that I will be addressing. We are in a relatively rich world in terms of inequality data – thanks in particlar to efforts of harmonized databases.

    But we lose much of the value – different how disconnected

     – that the rich doesn't translate to as  understanding

- Two worlds of inequality measurement

@@

    2) Hampered by the available data
        - **Coverage** – for instance no surveys.
        - **Quality and comarabilty of data** – Some secondary databases with global coverage mixes very heterogenous datapoints (WIID) (Find Atkinson and Brandolini ref from Handbook...). Those with a greater degree of harmonization tend to lack global coverage (LIS, EU-SILC, SEDLAC, until recently PIP)
        - **Some trade off between the two is inevitable** – There is no global survey of incomes. Countries run surveys based on their own needs and capacities. As such they differ. Saying something global means accepting a degree of heterogenity in the data. Similarly, definitions of income for tax purposes vary, impacting the comparability of tax data.

    3) Contradictory global views are presented by different global databases.
        - The data situation has improved drastically: thanks to harmonized databases with global coverage. In particular:
            - PIP – recently expanded to include high income countries
            - WID – Transition from WTID, and the DINA methodology, global coverage.
            - LIS: some data from middle and low income countries.
        - But the 'take home' messages from the institutions producing these databases offer very contradictory views.
            - "In recent decades, income inequality 
has increased in nearly all countries, 
but at different speeds, suggesting 
that institutions and policies matter in shaping inequality." (WIR 2018)
            - "[Across a range of approaches] average  inequality  within  countries  increased   in  the  1990s,  but  has  been  falling  in  recent years according to most specifications." (World Bank, 2016)

### 2023-02-10 PhD_pages/older_structure/data_appendices/3_A_prepare_KLEMS.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

---
title: "Appendix 3.A: Preparing EU KLEMS industry data"
format: html
execute:
  freeze: true  # never re-render during project render
---

@@

There are many (`r df %>% filter(VA == CAP & is.na(LAB)) %>% count() %>% as.numeric()`) country-year-industry observations where where CAP is 100% of VA, but LAB is missing.

Most of these (`r df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count() %>% as.numeric()`) are where both capital income and value added equal zero. Such observations appear only for the following industries:

@@

There is a similar set of country-year-industry observations (n = `r df %>% filter(VA == LAB & is.na(CAP)) %>% count() %>% as.numeric()`) where VA = LAB but CAP is NA. These we list here.

@@

There are some observations (n=`r main_industry_accounts %>% filter(is.na(VA) & is.na(CAP) & is.na(LAB)) %>% count() %>% as.numeric()`) without .... . We group these by country and industry.

@@

In general the sum across industries is very close to to the reported economy-wide aggregate. In the table below we show all observations where the ratio deviates from 100% by more than `r tolerance*100`%.

>>>>

---
title: "Appendix 3.A: Preparing EU KLEMS industry data"
format: html
execute:
  eval: false  # never run
---

@@

There are many (r df %>% filter(VA == CAP & is.na(LAB)) %>% count() %>% as.numeric()) country-year-industry observations where where CAP is 100% of VA, but LAB is missing.

Most of these (r df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count() %>% as.numeric()) are where both capital income and value added equal zero. Such observations appear only for the following industries:

@@

There is a similar set of country-year-industry observations (n = r df %>% filter(VA == LAB & is.na(CAP)) %>% count() %>% as.numeric()) where VA = LAB but CAP is NA. These we list here.

@@

There are some observations (n=r main_industry_accounts %>% filter(is.na(VA) & is.na(CAP) & is.na(LAB)) %>% count() %>% as.numeric()`) without .... . We group these by country and industry.

@@

In general the sum across industries is very close to to the reported economy-wide aggregate. In the table below we show all observations where the ratio deviates from 100% by more than r tolerance*100`%.

### 2023-02-10 PhD_pages/older_structure/data_appendices/3_B_prepare_OECD.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

---
title: "Appendix 3.B: Preparing OECD industry data"
format: html
execute:
  freeze: true  # never re-render during project render
---

@@

For instance, out of `r length(main_industry_codes)` aggregate industry codes, these country years have data for fewer than 18:

@@

In particular, the dataset contains only `r main_industry_accounts %>% filter(ind_code == "VU")%>%count()` country-year observations for `VU` (as compared to `r main_industry_accounts %>% filter(ind_code == "VA0")%>%count()` for `VA0`) and these are in general only NA or 0 values.

>>>>

---
title: "Appendix 3.B: Preparing OECD industry data"
format: html
execute:
  eval: false  # never run
---

@@

For instance, out of r length(main_industry_codes)` aggregate industry codes, these country years have data for fewer than 18:

@@

In particular, the dataset contains only r main_industry_accounts %>% filter(ind_code == "VU")%>%count()` country-year observations for `VU` (as compared to r main_industry_accounts %>% filter(ind_code == "VA0")%>%count()` for `VA0`) and these are in general only NA or 0 values.
